### 2021-09-15 README.Rmd
Add instructions for replicating

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)
```

# Causal inference in nonprofit management and public administration

<!-- badges: start -->
<!-- badges: end -->

### `r emoji_glue(":devil::bug:")`: Note on "corrupt caterpillar" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [**codename**](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change. The seed `1022` was generated with random atmospheric noise at [random.org](https://www.random.org/).

```{r generate-code-name}
library(codename)
codename_message()

codename(type = "ubuntu", seed = 1022)
```

## How to replicate

You can either [download the compendium as a ZIP file](/archive/main.zip) or use GitHub to clone or fork the compendium repository (see the green "Code" button at the top of the GitHub page).

We use the [**renv** package](https://rstudio.github.io/renv/articles/renv.html) to create a stable version-specific library of packages, and we use the [**targets** package](https://docs.ropensci.org/targets/) to manage all file dependencies and run the analysis. To reproduce the findings and re-run the analysis, do the following:

1. Download and install these fonts:
    - [Cochineal](https://fontesk.com/cochineal-typeface/)
    - [Econ Sans](https://github.com/hrbrmstr/hrbrthemes/tree/master/inst/fonts/econ-sans)
    - [Linux Libertine O](https://www.cufonfonts.com/font/linux-libertine-o) (also [here](https://sourceforge.net/projects/linuxlibertine/))
    - [Libertinus Math](https://github.com/alerque/libertinus)
    - [InconsolataGo](https://github.com/ryanoasis/nerd-fonts/tree/master/patched-fonts/InconsolataGo)
2. [Install R](https://cloud.r-project.org/) (and preferably [RStudio](https://www.rstudio.com/products/rstudio/download/#download)).
    - If you're using macOS, [install XQuartz too](https://www.xquartz.org/), so that you have access to the Cairo graphics library
3. Open `corrupt-caterpillar.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
4. Make sure you have a working installation of LaTeX:
    - *Easy-and-recommended way*: Install the [**tinytex** package](https://yihui.org/tinytex/) by running `install.packages("tinytex")` in the R console, then running `tinytex::install_tinytex()`
    - *Easy-but-requires-huge-4+-GB-download way*: Download TeX Live ([macOS](http://www.tug.org/mactex/); [Windows](https://miktex.org/))
5. If it's not installed already, R *should* try to install the **renv** package when you open the RStudio Project for the first time. If you don't see a message about package installation, install it yourself by running `install.packages("renv")` in the R console.
6. Run `renv::restore()` in the R console to install all the required packages for this project.
7. Run `targets::tar_make()` in the R console to automatically download all data files, process the data, run the analysis, and compile the paper and appendix.

Running `targets::tar_make()` will create several helpful outputs:

1. All project data in `data/`
2. An analysis notebook website in `analysis/_site/index.html`
3. PDF, HTML, and Word versions of the manuscript in `manuscript/output/`

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.