### 2024-11-09 manuscript/manuscript.qmd
Mega additions and edits and reorganization

---
title: "Clarifying Correlation and Causation: A Guide to Modern Quantitative Causal Inference in Nonprofit Studies"
short-title: "Clarifying Correlation and Causation"
date: 2024-11-08
author:
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  url: https://aysps.gsu.edu/phd-student/ye-meng/
  orcid: 0000-0002-6552-8338
correspondence-prefix: "Correspondence concerning this article should be addressed to"
published: Drafting
abstract: |
  Discovering causal relationships and testing theoretical mechanisms is a core endeavor of social science. Randomized experiments have long served as a gold standard for making valid causal inferences, but most of the data social scientists work with is observational and non-experimental. However, with newer methodological developments in economics, political science, epidemiology, and other disciplines, an increasing number of studies in social science make causal claims with observational data. As a newer interdisciplinary field, however, nonprofit studies has lagged behind other disciplines in its use of observational causal inference. In this paper, we present a hands-on introduction and guide to design-based observational causal inference methods. We first review and categorize all studies making causal claims in top nonprofit studies journals over the past decade to illustrate the field’s current of experimental and observational approaches to causal inference. We then introduce a framework for modeling and identifying causal processes using directed acyclic graphs (DAGs) and provide a walk-through of the assumptions and procedures for making inferences with a range of different methods, including matching, inverse probability weighting, difference-in-differences, regression discontinuity designs, and instrumental variables. We illustrate each approach with synthetic and empirical examples and provide sample R and Stata code for implementing these methods. We conclude by encouraging scholars and practitioners to make more careful and explicit causal claims in their observational empirical research, collectively developing and improving quantitative work in the broader field of nonprofit studies.
thanks: |
  Acknowledgments here if any
epigraph:
  - text: "Don't let us forget that the causes of human actions are usually immeasurably more complex and varied than our subsequent explanations of them."
    source: "Fyodor Dostoevsky, *The Idiot*, part IV, chapter 2"
keywords:
  - those
  - go
  - here
word-count: ""
bibliography: references.json
link-citations: true

params:
  wordcount: |
    <strong>{{< words-sum body-ref >}} total words</strong>: {{< words-body >}} in the body • {{< words-ref >}} in the references
---

::: {.content-visible when-meta="wordcount-banner"}
{{< include _extensions/andrewheiss/wordcount/banner.html >}}
:::

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(
  echo = FALSE, include = FALSE, 
  warning = FALSE, message = FALSE
)

knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "100%", collapse = TRUE
)

options(
  digits = 3, width = 120
)
```

```{r libraries-data, include=FALSE}
library(tidyverse)
library(patchwork)
library(ggdag)
library(dagitty)
library(targets)
library(here)

# Targets stuff
tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

# Plotting functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
```

```{=latex}
% xelatex + unicode-math breaks over/underbrackets so we need to save the 
% bracket environments and then restore them after \begin{document}. 
% See https://github.com/latex3/mathtools/issues/29

\makeatletter
\@ifpackageloaded{unicode-math}{
% Restore mathtools' brackets
\let\underbracket=\normalunderbracket
\let\overbracket=\normaloverbracket
}{}
\makeatother
```

::: {.content-visible when-format="html"}
# Introduction {.hide}
:::

Causal inference important in social sciences, but hard. Nobel prize for Angrist and Imbens and gang...

Answering causal questions is an important endeavor for nonprofit researchers and practitioners. Nonprofits serve social goals and create and implement a range of programs to accomplish those goals. Measuring the impact of nonprofit programming is essential both for grantors that require evidence of program impact and for nonprofit program managers who are interested in whether their programs lead to the desired outcomes.

Causal work in the nonprofit policy and management literature has increased in the past decade (see fig1), but it remains a less explored topic. The nonprofit literature is full of careful quantitative and statistical work, explicitly descriptive, implicitly causal (with weasel words), and explicitly causal.

In this paper, we provide a practical reference, guide, and review of modern causal inference methods. For the sake of accessibility, we provide sample data, R and Stata code, minimal—and extensively annotated—mathematical notation. We offer a unified, cross-disciplinary framework for thinking about causal identification, spanning econometrics and epidemiology, providing a toolbox for nonprofit researchers to think more carefully about causally-oriented research design and analysis, leading to more effective evaluations of programs, policies, and interventions.

Roadmap—in this paper we blah blah blah

---

Stuff from first version of introduction:

Causal questions are at the root of many questions in nonprofit studies. Do transparency and accountability improve nonprofit performance? Does being listed on a transparency-focused website like GuideStar increase donations? Other things

Researchers can investigate causal mechanisms using many different methods. Qualitative scholars can do focus groups, interviews, constructivist stuff, case studies, whatever—deep dives into specific mechanisms in particular cases. In the age of large, publicly available datasets—as well as nonprofits, foundations, and government agencies swimming in data—quantitative research can interrogate causal mechanisms as well. But at the root of all this data lies the statistical adage “correlation does not imply causation”. In the absence of experimental data, where researchers can randomly assign nonprofits or donors to treatment and control conditions to eliminate any confounding differences between the two groups, making causal claims from quantitative studies is difficult. It is not impossible though. Advances in econometric, epidemiological, statistical, and computer scientific methods over the past few decades have led to a “credibility revolution,” where researchers can use carefully constructed research designs to estimate causal effects using observational data

In this paper, we provide an introductory primer to thinking about causal questions using observational data. This will allow nonprofit researchers to make more careful causal claims…

The "causal revolution" / "credibility revolution" has been happening for the past decade in social science disciplines - important because of reasons

How has the discipline of nonprofit studies embraced the causal revolution?

Are there fears of the word "causation" and an emphasis on pure association?

Nonprofit data is very observational - quantitative data from past X years in N journals relies on experimental data Y% of the time, and observational data Z% of the time.

Similar to recent work by @BaBerrettCoupet:2021 and @MaEbeiddeWit:2023 and @Rohrer:2018, in this paper, we provide a introductory primer / framework for approaching causal questions with observational data - accessible and approachable guide to nonprofit researchers and practitioners. The framework here is crucial for general academic research, policy analysis, and program evaluation, among other common types of nonprofit research.

Importantly, in this article with are concerned only with *quantitative* causal inference, or the use of statistical approaches to identify and isolate causal mechanisms. We do not assert that quantitative methods are the only avenue for answering causal questions. Qualitative and mixed methods like focus groups, action research, ___, ___, and ___ are designed to explore causal mechanisms with rich detail (@cite1). Some qualitative methods are at epistemological odds with the whole endeavor of causal effects (@cite2).

Instead, we address the perennial caveat taught in introductory statistics classes: that correlation does not imply causation. What *does* imply causation? How can we use statistical tools to explore causal questions using observational data?

---

# Why care about causation in nonprofit studies? (MENG)

Graph and analysis of causal stuff in top 3 here

[@Samii:2016]

> 941 - "Causal empiricism is associated with “identification strategy” research designs."

> This is different from quantitative “pseudo-general pseudo-facts” that come from multiple regression

>> In quantitative research, pseudo-facts are statistical results that are interpreted erroneously in terms of their causal implications, and pseudo-general findings are ones that are erroneously described as applying to a more general class of units than is immediately warranted.

Don't make unwarranted causal-ish statements, but also don't be afraid of the "c-word" [@Hernan:2018]

Language can imply causality even if authors explicitly eschew causal language and identification strategies [@HaberWietenRohrer:2021]

Weasel words - association, determinants, etc. 
Determinants and prediction is fine! The focus is on getting the most accurate prediction of the outcome
When making causal claims, though, the focus is on one of the Xs—one lever that a government agency or nonprofit organization can manipulate to affect some sort of change

---

Textbook treatments of causal inference approahces are typically abstract and generic, explaining how treatment $X$ causes outcome $Y$ after adjusting for confounders $Z$. To make these principles as concrete as possible, we use a practical running example relevant to nonprofit education and practice. Experiential philanthropy - Newark, BYU Grantwell and Be a Philanthropist. These programs provide students interested in the nonprofit sector with hands on, practical experience with philanthropy. Beyond their pedagogical aims, these programs are designed to instill a long-term interest and participation in philanthropy.

> Do experiential philanthropy programs cause long-term charitable or prosocial behavior?

We do not provide any evidence to answer this question directly. Rather, we use simulated data to explore multiple hypothetical research designs that could be used to quantititavely measure the causal impact of these programs. These causal inference approaches can be applied to a wide range of nonprofit-focued questions, though some methods are better suited than others, depending on the context of the research question. TODO We provide examples of possible nonprofit-related causal inquiries that are amenable to each method

# A visual vocabulary for causal inference

## Causation through listening

Judea Pearl, computer science, and epidemiology blah blah blah provide us with a specialized framework for describing causal theories and making causal inferences.

At the core of causal inquiry is the notion of "causation," or what it means when we say that an intervention causes an outcome. In Pearl's framework, causation can be defined using a metaphor of listening and responding:

> A variable $X$ is a cause of a variable $Y$ if $Y$ in any way relies on $X$ for its value.… $X$ is a cause of $Y$ if $Y$ listens to $X$ and decides its value in response to what it hears. [@PearlGlymourJewell:2016, 5--6].

In the context of our example causal question, we posit that prosocial behavior listens to, responds to, or is caused by experiential philanthropy. We have a theoretical reason to believe that the intervention and the outcome are associated with each other in a sequential, nonspurious way—it is reasonable to assume that having hands-on experience with philanthropy today would influence future attitudes toward charitable giving. Specific educational interventions are not the only cause of prosocial behavior. This behavior listens and responds to other factors, including family background, income, education, and personal attitudes and opinions towards volunteerism and charity—experiential philanthropy is just one of many causes.

## Directed acyclic graphs (DAGs)

We can formally represent our theory of how an intervention ($X$) causes an outcome ($Y$) using a directed acyclic graph, or DAG [@PearlMackenzie:2018; @PearlGlymourJewell:2016; @MorganWinship:2014]. DAGs encode our understanding of the data generating process, or which phenomena cause the treatment, the outcome, and both simultaneously. These causal graphs are a philosophical model of the associations between different phenomena and our assumptions behind those relationships.

- **Nodes** - nodes can be unmeasurable, or even unobserved
- **Edges** - Arrows indicate a relationship, or the passing of *statistical association* between nodes - intervening on one node leads to changes in another node - When two nodes are connected by an arrow, we are stating that there is an assumed causal relationship between those nodes; when there is no arrow between nodes, we are explicitly stating that there is no relationship between the two. Unlike other graphical approaches like structural equation models (SEMs) that assume all relationships are additive and linear, DAG arrows are nonparametric and can represent any kind of functional form (linear, multiplicative, polynomial, exponential, etc.). Accordingly, when we say that $X \rightarrow Y$, we mean that $X$ causally affects $Y$ in some general way, not necessarily linearly [@Rohrer:2018].
- **Acyclicity** - can't get back to a node - represents flow of time, or temporal ordering - if things are cyclical, like hiring new staff $\leftrightarrow$ increased capacity, you can make these nodes time-based: hiring new staff_t → increased capacity_t → hiring new staff_t+1 → increased capacity_t+1

```{r}
#| label: fig-dag-basic-real
#| fig-width: 4.5
#| fig-height: 2.75
#| out-width: 60%
#| fig-cap: "An example directed acyclic graph that shows how experiential philanthropy causes charitable behavior, while incomes causes both"
#| include: true

data_confounder_example <- dagify(
  Y ~ Z + X,
  X ~ Z,
  coords = list(x = c(X = 1, Y = 3, Z = 2),
                y = c(X = 1, Y = 1, Z = 2)),
  labels = c(X = "Experiential\nphilanthropy", Y = "Charitable\nbehavior", Z = "Income")
) |> 
  tidy_dagitty()

ggplot(data_confounder_example, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(edge_width = 1) +
  geom_dag_point(aes(color = name), size = 12) +
  geom_dag_text(aes(label = name)) +
  geom_dag_text(
    data = filter(data_confounder_example, name != "Z"), aes(label = label), 
    size = 4, nudge_y = -0.32, color = "black", lineheight = 1
  ) +
  geom_dag_text(
    data = filter(data_confounder_example, name == "Z"), aes(label = label), 
    size = 4, nudge_y = 0.25, color = "black", lineheight = 1
  ) +
  scale_color_manual(values = c("X" = clrs$green, "Y" = clrs$blue, "Z" = clrs$yellow), guide = "none") +
  coord_cartesian(xlim = c(0.75, 3.25), ylim = c(0.6, 2.2)) +
  theme_np_dag()
```

Importantly, DAGs do more than simply represent the relationships and associations between variables. They provide us with ways to isolate specific relationships of interest. TODO: Rohrer, 28: Even if a researcher does not rely fully on a DAG, mapping out the underlying data-generating process

## Statistical associations and causal structures

General explanation of DAGs and confounding/colliding/mediating and d-separation and do-calculus, etc.

```{r}
#| label: fig-dag-associations
#| fig-width: 9
#| fig-height: 2.75
#| fig-cap: "Three types of relationships in DAGs"
#| include: true

data_confounder <- dagify(
  Y ~ Z + X,
  X ~ Z,
  coords = list(x = c(X = 1, Y = 3, Z = 2),
                y = c(X = 1, Y = 1, Z = 2))
) |> 
  tidy_dagitty() |> 
  mutate(arrow_color = ifelse(name == "Z", clrs$orange, "grey80")) |> 
  mutate(node_color = ifelse(name == "Z", TRUE, FALSE))

confounder <- ggplot(data_confounder, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(aes(edge_colour = arrow_color), edge_width = 1) +
  geom_dag_point(aes(color = node_color), size = 12) +
  geom_dag_text(data = filter(data_confounder, name != "Z"), size = 4) +
  geom_dag_text(data = filter(data_confounder, name == "Z"), color = "white", size = 4) +
  scale_color_manual(values = c("grey80", clrs$orange), guide = "none") +
  coord_cartesian(ylim = c(0.95, 2.05)) +
  labs(title = "Confounder") +
  theme_np_dag() +
  theme(plot.title = element_text(hjust = 0.5, color = clrs$orange))

data_mediator <- dagify(
  Y ~ X + Z,
  Z ~ X,
  coords = list(x = c(X = 1, Y = 3, Z = 2),
                y = c(X = 1, Y = 1, Z = 2))
) |> 
  tidy_dagitty() |> 
  mutate(arrow_color = case_when(
    name == "Z" & to == "Y" ~ clrs$purple, 
    name == "X" & to == "Z" ~ clrs$purple,
    TRUE ~ "grey80")) |> 
  mutate(node_color = ifelse(name == "Z", TRUE, FALSE))

mediator <- ggplot(data_mediator, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(aes(edge_colour = arrow_color), edge_width = 1) +
  geom_dag_point(aes(color = node_color), size = 12) +
  geom_dag_text(data = filter(data_mediator, name != "Z"), size = 4) +
  geom_dag_text(data = filter(data_mediator, name == "Z"), color = "white", size = 4) +
  scale_color_manual(values = c("grey80", clrs$purple), guide = "none") +
  coord_cartesian(ylim = c(0.95, 2.05)) +
  labs(title = "Mediator") +
  theme_np_dag() +
  theme(plot.title = element_text(hjust = 0.5, color = clrs$purple))

data_collider <- dagify(
  Y ~ X,
  Z ~ X + Y,
  coords = list(x = c(X = 1, Y = 3, Z = 2),
                y = c(X = 1, Y = 1, Z = 2))
) |> 
  tidy_dagitty() |> 
  mutate(arrow_color = ifelse(to == "Z", clrs$red, "grey80")) |> 
  mutate(node_color = ifelse(name == "Z", TRUE, FALSE))

collider <- ggplot(data_collider, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(aes(edge_colour = arrow_color), edge_width = 1) +
  geom_dag_point(aes(color = node_color), size = 12) +
  geom_dag_text(data = filter(data_collider, name != "Z"), size = 4) +
  geom_dag_text(data = filter(data_collider, name == "Z"), color = "white", size = 4) +
  scale_color_manual(values = c("grey80", clrs$red), guide = "none") +
  coord_cartesian(ylim = c(0.95, 2.05)) +
  labs(title = "Collider") +
  theme_np_dag() +
  theme(plot.title = element_text(hjust = 0.5, color = clrs$red))

combined_associations <- (confounder | plot_spacer() | mediator | plot_spacer() | collider) +
  plot_layout(widths = c(0.32, 0.02, 0.32, 0.02, 0.32))
combined_associations
```

This stuff in @fig-dag-associations: (@Elwert:2013 for forks, chains, inverted forks terminology)

- Confounding - forks
- Mediation - chains
- Collision - inverted forks [@ElwertWinship:2014; @KnoxLoweMummolo:2020]

Confounding is a major scary issue though - especially if it's unmeasured or unobservable…

## Bad controls

Even if we do not rely on a DAG for complete identification of causal effects, mapping out the relationships between nodes in the data-generating process is a useful exercise in any kind of quantitative, regression-based analysis, including non-causal descriptive work. Researchers often take a "kitchen sink" approach to statistical control, including all potential covariates in a model. Understanding the underlying DAG provides clearer guidance about what to control for, helping researchers avoid "bad controls" [@cite]. For instance, if a potential covariate is only associated with the outcome and has no relationship to the treatment, it is not a confounder and need not be included as a control variable in a model, since it can reduce the precision of the model estimates. If a potential covariate is a collider, caused by both the treatment and the outcome, the model results will be biased or distorted.

In addition to identification, DAGs provide additional statistical insight and guidance regarding control variables or covariates in regression models

Bad controls and colliders

Good/bad controls: <https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3689437> - <https://twitter.com/analisereal/status/1512596580632707078>

Table 2 fallacy + @KeeleStevensonElwert:2020 thing

Post-treatment control bias - cite that one paper by Brendan Nyhan - DAGs make it obvious which nodes are post-treatment, since they appear after the treatment node in the causal chain in the graph

# Identifying and isolating causal relationships

## Fundamental problem of causal inference

Formulating and testing hypotheses is the foundation of scientific inquiry. In chemistry, researchers can intervene with blah blah H2O

People and individuals, however, do not respond in identical ways. Counterfactuals - what would have happened in the absense of an intervention to the same person or organization.

Rubin notation for individual potential outcomes
Treatment effect = whatever

However, it is impossible to observe or measure both Yi0 and Yi1. A charitable foundation interested in poverty reduction cannot give a grant to a nonprofit and simultaneously not give it to the same nonprofit in order to measure the impact of its donation. What the recipient nonprofit receives (or does not receive) from the foundation becomes a realized outcome rather than a potential outcome, and the counterfactual outcome is forever unrealized and unmeasurable.

To get around this, we can take the average (or expectation, often indicated with the mathematical function E) of many units that receive the treatment and compare it with the average of comparable units that don’t receive the treatment

Rubin notation: Potential outcomes notation

To help clarify that blah blah, we'll introduce one more piece of mathematical notation - the do() operator

Pearl notation: $\operatorname{do}(x)$ notation = an intervention.

## Selection bias

In addition to the fact that potential outcomes are unobservable, one more characteristic of observational data makes causal inference more complicated.

Individuals and organizations have agency and choose their interventions

A foundation interested in reducing poverty will typically research a range of candidate nonprofits and select the one that fits their own internal criteria. Nonprofits applying for a grant from that foundation will tailor their applications to meet the foundation's preferences. Accordingly, there are systematic differences between nonprofits that receive a grant and those that don't. It is tempting to measure the average outcome of recipient nonprofits and compare it with the average outcome of non-recipient nonprofits, but this estimate would be incorrect.

Formula showing wrong effect

## Causal identification and *d*-separation

To identify a causal relationship in a DAG, we have to ensure that the arrow between the treatment and outcome nodes is isolated and that the treatment and outcome nodes are not linked through any other pathways

In addition to encoding our philosophy and theory of the data generating process, DAGs also serve as an important statistical tool for isolating or identifying causal quantities of interest. Identification strategy definition:

> The central role of an identification strategy is to provide a logic for establishing that D is independent of potential values of Y, thus allowing the analyst to interpret observed associations as causal effects. [@KeeleStevensonElwert:2020]

- d-connection and d-separation - statistical association cannot pass between nodes, either because the arrows are drawn in a way that makes it so information does not link the two, or because conditioning/adjustment blocks the pathway

## *do*-calculus and statistical adjustment

In experiments, the researcher has total control over assignment to treatment, which means all edges/arrows that might influence treatment can be removed. There is no confounding to worry about and we can measure the exact causal effect of X on Y. In potential outcomes language, you still can't see each individual's yes and no response, but you can average all the yeses and noes and get an average causal effect

With observational data, we'd like to measure $\mathbf{E}(y \mid \operatorname{do}(x))$ but we can only actually see and measure $\mathbf{E}(y \mid x)$, and as shown in @eq-correlation-not-causation, these two expressions are not the same. This is a formal statement of the phrase "correlation isn't causation":

$$
{\color{gray} \overbrace{{\color{orange} \underbracket[0.25pt]{{\color{black} \mathbf{E}(y \mid \operatorname{do}(x)) \vphantom{\frac{1}{2}}}}_{\color{orange} \text{``Causation"}}}}^{\color{gray} \mathclap{\substack{\text{The average} \\ \text{population-level} \\ \text{change in $y$ when} \\ \textit{directly intervening} \\ \text{(or doing) $x$}}}}}
\quad \neq \quad
{\color{gray} \overbrace{\color{purple} \underbracket[0.25pt]{{\color{black} \mathbf{E}(y \mid x)} \vphantom{\frac{1}{2}}}_{\color{purple} \text{``Correlation"}}}^{\color{gray} \mathclap{\substack{\text{The average} \\ \text{population-level} \\ \text{change in $y$ when} \\ \text{accounting for} \\ \textit{observed } x}}}}
$$ {#eq-correlation-not-causation}

What we want to be able to do is transform the $\mathbf{E}(y \mid \operatorname{do}(x))$ expression into something without the $\operatorname{do}(x)$, or something *do*-free. A set of three systematic rules for analyzing and decomposing causal graphs known as *do*-calculus provide certain conditions under which we can treat an interventional $\operatorname{do}(\cdot)$ expression like an observed value instead. A complete exploration of these three rules of *do*-calculus go beyond the scope of this paper, but lots of resources like Pearl, that one textbook, other things in my blog post, etc. [@Pearl:2012; @Pearl:2019]

The most common derivation of the rules of *do*-calculus is an approach called "backdoor adjustment". By adjusting or controlling for nodes that open up backdoor paths between the treatment and outcome nodes, we can mathematically transform a $\operatorname{do}(\cdot)$ expression into something based solely on observational data. Formally, the backdoor adjustment formula is defined in @eq-backdoor:

$$
{\color{gray} \overbrace{\color{black} \mathbf{E}(y \mid \operatorname{do}(x)) \vphantom{\frac{1}{2}}}^{\substack{\text{Causal effect} \\ \text{of $x$ on $y$}}}}
\quad=\quad
{\color{gray} \underbrace{{\color{black} \sum_z}}_{\mathclap{\substack{\text{Sum across} \\ \text{all values of $z$}}}}}
{\color{gray} \overbrace{\color{black} \mathbf{E} (y \mid x, z) \vphantom{\frac{1}{2}}}^{\mathclap{\substack{\text{Conditional} \\ \text{mean of $y$,} \\ \text{given $x$ and $z\dots$}}}}}
\enspace\times\enspace
{\color{gray} \overbrace{\color{black} \mathbf{P}(z) \vphantom{\frac{1}{2}}}^{\mathclap{\substack{\text{$\dots$ weighted} \\ \text{by $z$}}}}}
$$ {#eq-backdoor}

Put more simply, \@ref(eq:backdoor) demonstrates that we can remove the interventional $\operatorname{do}(x)$ from the left-hand side of the equation by controlling for (or conditioning on) all the confounders $z$ that open up a backdoor pathway between treatment and outcome. As a simplified illustration, suppose that the relationship between treatment and outcome is confounded only by a nonprofit's size, which is measured as either large or small. Applying this backdoor adjustment formula would entail finding average value of the outcome conditioned on the treatment among large nonprofits, multiplied by the proportion of large nonprofits, added to the average value of the outcome conditioned on the treatment among small nonprofits, multiplied by the proportion of small nonnprofits. The resulting sum would then be the unconfounded causal effect.

In practice, statistical adjustment rarely involves a single binary confounder. For instance, in the causal graph in Fig X, X, Y, and Z all open up backdoors between treatment and outcome, and all three would need to be adjusted for. We will provide a practical demonstration of more common adjustment strategies when there are multiple confounders in section X. At this point, what is important to note is that adjusting for confounding nodes allows us to isolate the single pathway between treatment and outcome. Because spurious statistical associations from other nodes have been blocked statistically, the relationship we care about is identified and we can talk about the *causal* effect of the treatment on the outcome.

TODO: Plot of backdoor and frontdoor adjustment DAGs, but using nonprofit situations

A less common derivation of the rules of *do*-calculus is frontdoor adjustment, commonly used when confounding is unobserved and undertheorized and unmeasurable. Smoking genetics tar cancer thing - Bellemare paper example [@BellemareBloemWexler:2020] - frontdoor adjustment formula here? We do not provide a complete example here—see Bellemare for that—but again, what is most important here is that we can again mathematically transform a quantity with an interventional do(x) into a do-free quantity, meaning that we can make causal claims from observational data.

The backdoor and frontdoor criteria are the most common applications of *do*-calculus because they are readily apparent in causal graphs—it is possible to see forks joining exposure and outcome and identify backdoors, or see measurable mediating nodes that could be used as front doors. In more complex DAGs, these backdoor and frontdoor shortcuts might not be readily visible. In that case, there are software packages that algorithmically work through the various rules of *do*-calculus to determine the set of nodes that need to be adjusted in order to isolate the x → y relationship. Not every DAG is identifiable; but any identifiable DAG can be identified.

---

The logic of *do*-calculus tells us what nodes or variables been to be adjusted for to isolate the treatment → outcome arrow, but the DAG provides no guidance about how to actually make these adjustments.

There are two general approaches for doing this:

- Circumstantial identification
- Adjustment-based identification

## Circumstantial identification

[@AngristPischke:2009; @AngristPischke:2015]

The language of causal graphs, identification, and adjustment provide a universal grammar for discussing causal effects. Commonly used approaches in econometrics and other social science disciplines can be written as causal graphs [see @fig-design-based-dags]

```{r}
#| label: fig-design-based-dags
#| fig-width: 9
#| fig-height: 6
#| fig-cap: "Possible DAGs for common design-based experimental and quasi-experimental approaches to causal inference. Red arrows represent the identified and isolated relationship between treatment $x$ and outcome $y$. Square nodes represent statistical adjustment."
#| include: true

data_rct <- dagify(
  Y ~ Z + X,
  coords = list(x = c(X = 1, Y = 3, Z = 2),
                y = c(X = 1, Y = 1, Z = 2)),
  exposure = "X",
  outcome = "Y"
) |> 
  tidy_dagitty() |> 
  node_status() |> 
  mutate(arrow_color = ifelse(name == "X", clrs$red, "grey80"))

rct <- ggplot(data_rct, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(aes(edge_colour = arrow_color), edge_width = 1) +
  geom_dag_point(aes(color = status), size = 12) +
  geom_dag_text(size = 4) +
  annotate(geom = "text", x = 1, y = 1.2, label = "X = x") +
  scale_color_manual(values = c(clrs$green, clrs$blue), 
                     na.value = clrs$yellow, guide = "none") +
  coord_cartesian(ylim = c(0.95, 2.05)) +
  labs(title = "Randomization", 
       subtitle = "Randomization deletes all arrows into X") +
  theme_np_dag() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

data_iv <- dagify(Y ~ X + U,
                  X ~ I + U,
                  exposure = "X",
                  outcome = "Y",
                  latent = "U",
                  coords = list(x = c(X = 1, Y = 3, U = 2, I = 0),
                                y = c(X = 1, Y = 1, U = 2, I = 1))) |> 
  tidy_dagitty() |> 
  adjust_for(var = "I") |> 
  node_status() |> 
  mutate(arrow_color = ifelse(name == "X", clrs$red, "grey80"))

iv <- ggplot(data_iv, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(aes(edge_colour = arrow_color), edge_width = 1) +
  geom_dag_point(aes(color = status, shape = adjusted), size = 12) +
  geom_dag_text(size = 4) +
  scale_color_manual(values = c(clrs$green, "grey80", clrs$blue), 
                     na.value = clrs$yellow, guide = "none") +
  scale_shape_manual(values = c(15, 19), guide = "none") +
  coord_cartesian(ylim = c(0.95, 2.05)) +
  labs(title = "Instrumental variable", 
       subtitle = "Find effect of instrument (I) on X, then find effect of (X | I) on Y") +
  theme_np_dag() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

data_diff_diff <- dagify(Y ~ X + Time + Location,
                         X ~ Time + Location,
                         exposure = "X",
                         outcome = "Y",
                         coords = list(x = c(X = 1, Y = 3, Time = 2, Location = 2),
                                       y = c(X = 1, Y = 1, Time = 2, Location = 0))) |> 
  tidy_dagitty() |> 
  adjust_for(var = c("Time", "Location")) |> 
  node_status() |> 
  mutate(arrow_color = ifelse(name == "X", clrs$red, "grey80"))

diff_diff <- ggplot(data_diff_diff, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(aes(edge_colour = arrow_color), edge_width = 1) +
  geom_dag_point(aes(color = status, shape = adjusted), size = 12) +
  geom_dag_text(data = filter(data_diff_diff, name %in% c("X", "Y")), size = 4) +
  geom_dag_text(data = filter(data_diff_diff, name == "Location"), 
                nudge_y = 0.37, size = 4) +
  geom_dag_text(data = filter(data_diff_diff, name == "Time"), 
                nudge_y = -0.37, size = 4) +
  scale_color_manual(values = c(clrs$green, clrs$blue), 
                     na.value = clrs$yellow, guide = "none") +
  scale_shape_manual(values = c(15, 19), guide = "none") +
  coord_cartesian(ylim = c(-0.05, 2.05)) +
  labs(title = "Difference-in-differences", 
       subtitle = "Adjust for both time (e.g., year) and location (e.g., country, state)") +
  theme_np_dag() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

data_rdd <- dagify(Y ~ X + Running + U,
                   X ~ Threshold + U,
                   Running ~ U,
                   Threshold ~ Running,
                   exposure = "X",
                   outcome = "Y",
                   latent = "U",
                   coords = list(x = c(X = 1, Y = 4, Running = 3, Threshold = 2, U = 3),
                                 y = c(X = 1, Y = 1, Running = 2, Threshold = 1.75, U = 3))) |> 
  tidy_dagitty() |> 
  adjust_for(var = c("Threshold", "Running")) |> 
  node_status() |> 
  mutate(arrow_color = ifelse(name == "X", clrs$red, "grey80"),
         name = recode(name, "Running" = "Running\nvariable"),
         linetype = ifelse(name == "U" & to == "X", "21", "solid"))

rdd <- ggplot(data_rdd, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(aes(edge_colour = arrow_color, edge_linetype = linetype), edge_width = 1) +
  geom_dag_point(aes(color = status, shape = adjusted), size = 12) +
  geom_dag_text(data = filter(data_rdd, name %in% c("X", "Y", "U")), size = 4) +
  geom_dag_text(data = filter(data_rdd, name == "Running\nvariable"),
                nudge_y = -0.47, size = 4, lineheight = 1) +
  geom_dag_text(data = filter(data_rdd, name == "Threshold"),
                nudge_y = -0.35, size = 4) +
  scale_color_manual(values = c(clrs$green, "grey80", clrs$blue), 
                     na.value = clrs$yellow, guide = "none") +
  scale_shape_manual(values = c(15, 19), guide = "none") +
  coord_cartesian(ylim = c(0.95, 3.05)) +
  labs(title = "Regression discontinuity", 
       subtitle = "Adjust for both the running variable and the threshold") +
  theme_np_dag() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

plot_econometrics <- ((rct | iv) / plot_spacer() / (diff_diff | rdd)) +
  plot_layout(heights = c(0.49, 0.02, 0.49))
plot_econometrics
```

### RCTs

No need to control for a ton of things in an RCT precisely because the arrows into X get deleted. No need to worry about perfect balance checks because the researcher has control over and understands the data generating process and assignment to treatment. Look at that one knitted Rmd on RCT FAQs: https://macartan.github.io/i/notes/rct_faqs.html - CONSORT also says to stop doing balance tests - only really need to control for things that might be predictive (https://twitter.com/statsepi/status/1115902270888128514?s=21), but theoretically anything that influences the allocation to treatment is taken care of by randomization.

- Explanation + DAG
- Illustration of how to use it
- Review of existing nonprofit studies that use it - survey vignette experiments, conjoint experiments, field experiments, other kinds of RCTs
- Possible nonprofit research questions that could use it

### Diff-in-diff + TWFE

Time / location, TWFE stuff

- Explanation + DAG
- Illustration of how to use it
- Review of existing nonprofit studies that use it
- Possible nonprofit research questions that could use it

### RDD

Threshold/cutpoint

Adjusting for the threshold and only looking at data right around it makes it so that we can treat the sample as if it were random (by assumption), which then means we can delete any arrows going into X just like an RCT

Cite Nick's *The Effect* - refer to his website with the DAGs and animations

- Explanation + DAG
- Illustration of how to use it
- Review of existing nonprofit studies that use it
- Possible nonprofit research questions that could use it

### IV

IVs have to meet the exclusion restriction - the instrument can only influence the outcome through the treatment. DAGs make this assumption very clear. There cannot be an arrow connecting the instrument to the outcome. DAGs also inform the exogeneity assumption—no other nodes in the graph can feed into the instrument node

It's like frontdoor adjustment (https://www.stat.cmu.edu/~cshalizi/402/lectures/23-causal-estimation/lecture-23.pdf) - Really we're finding the causal effect of I on X, then X on Y, generally through 2SLS

These used to be common, but have become less popular because of the difficulty in finding a valid instrument that meets all the criteria for clean identification. @rainfall-instrument-paper. Randomized promotion works well, though - @worldbank-book

- Explanation + DAG
- Illustration of how to use it
- Review of existing nonprofit studies that use it
- Possible nonprofit research questions that could use it

## Adjustment-based identification

### Regression adjustment

- Explanation + DAG
- Illustration of how to use it
- Review of existing nonprofit studies that use it
- Possible nonprofit research questions that could use it

### IPW and g-computation

- Explanation + DAG
- Illustration of how to use it
- Review of existing nonprofit studies that use it
- Possible nonprofit research questions that could use it

Matching creates entirely new treatment/control populations

IPW creates comparable pseudo populations

MAYBE: Make an image with little shaded people showing how matching and IPW work, similar to Torres:2020 (maybe with https://github.com/propublica/weepeople ?)

---

Despite its popularity in epidemiology and public health, to our knowledge, there are no studies in NVSQ, Voluntas, or NML that employ inverse probability weighting for covariate adjustment for causal inference. There are X that use matching, and there are many that use regression-based adjustment in explicitly non-causal ways (but Haber et al paper finds that these things imply causation anyway). For instance, Huafang et al. paper explicity talks about associations after controlling for ostensible confounders, but also uses causal language like "program impacts" - C-word article and not shying away from causation

## Hybrid approaches

Briefly describe combinations of the two approaches, like synthetic controls, DiD with matching, RDD with covariate adjustment, etc.

Briefly mention other methods, like marginal structural models [@BlackwellGlynn:2018]?

# Tools for researchers and practitioners

Code in R and Stata? Point to other resources like The Mixtape, The Effect, Pearl's stuff, Morgan and Winship?

[@Huntington-Klein:2021; @Cunningham:2021]

# Conclusion

A call for more causation / more careful thinking around causation. Prediction is fine. But don't automatically run away from causal work.

We believe that adjustment-based approaches like regression adjustment, matching, and IPW are highly amenable to nonprofit administrative data. Finding specific situations like diff-in-diff is popular in public policy analysis and econometrics because the scale of the data is much larger—researchers can analyze the effect of state- or national-level policies in difference-in-difference designs, or observe the behavior of millions of participants in programs like Medicaid or the Affordable Care Act for regression discontinuity designs. A smaller nonprofit interested in evaluating the impact of a new community program is most likely unable to consider these types of situation-specific quasi-experimental designs to isolate the causal effect. However, with careful theoretical thinking; the development of a robust DAG that describes potential confounders, mediators, and colliders; the use of high quality administrative data; and intentional sensitivity analysis that probes the strength of possible causal relationships when faced with unmeasured confounding [@that-sensmakr-paper; @lucys-tipr-paper]; nonprofit researchers can tell plausible causal stories.

# References