### 2024-11-21 methods/rct.qmd

---
title: "RCT"
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "95%",
  warning = FALSE,
  collapse = TRUE,
  cache.extra = 1234  # Change number to invalidate cache
)

options(
  digits = 3, width = 120,
  tinytable_tt_digits = 2
)
```

```{r}
#| label: packages-data
#| warning: false
#| message: false

library(tidyverse)
library(targets)
library(modelsummary)
library(tinytable)
library(ggdist)
library(scales)
library(ggdag)
library(dagitty)
library(parameters)

tar_config_set(
  store = here::here('_targets'),
  script = here::here('_targets.R')
)

rct <- tar_read(data_rct)

# Plotting functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
set_annotation_fonts()

gof_map <- tribble(
  ~raw,            ~clean,               ~fmt, ~omit,
  "nobs",          "N",                  0,    FALSE,
  "r.squared",     "$R^2$",          2,    FALSE,
  "adj.r.squared", "\\(R^2\\) adjusted", 2,    FALSE
)
```

## DAG

```{r}
#| label: plot-dag-rct
#| fig-width: 6
#| fig-height: 6

coords <- tribble(
  ~name, ~x,   ~y,   ~type,         ~color,         ~label,
  "X",   1,    3,    "Treatment",   clrs$Prism[4],  "Experiential\nphilanthropy",
  "Y",   3,    3,    "Outcome",     clrs$Prism[2],  "Charitable\nbehavior",
  "Z1",  5/3,  4.2,  "Confounder",  clrs$Prism[6],  "Social awareness",
  "Z2",  7/3,  4.2,  "Confounder",  clrs$Prism[6],  "Age",
  "Z3",  5/3,  5,    "Confounder",  clrs$Prism[6],  "GPA",
  "Z4",  7/3,  5,    "Confounder",  clrs$Prism[6],  "Income",
  "M",   2,    2,    "Mediator",    clrs$Prism[3],  "Community\nconnections",
  "C",   2,    1,    "Collider",    clrs$Prism[8],  "Employment in\nnonprofit sector"
) |> 
  mutate(type = fct_inorder(type))

data_full_dag <- dagify(
  Y ~ X + Z1 + Z2 + Z3 + Z4 + M,
  M ~ X,
  C ~ X + Y,
  coords = coords,
  labels = pull(coords, label, name = name)
) |> 
  tidy_dagitty() |> 
  left_join(select(coords, name, type, color), by = join_by(name)) |> 
  mutate(arrow_color = case_when(
    name == "X" & to == "Y" ~ "black",
    to == "C" ~ clrs$Prism[8],
    to == "M" | (name == "M" & to == "Y") ~ clrs$Prism[3],
    .default = color
  ))

ggplot(data_full_dag, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(edge_width = 0.5) +
  geom_dag_edges(aes(edge_color = arrow_color), edge_width = 0.5) +
  geom_dag_point(aes(color = type), size = 12) +
  geom_dag_label(
    data = filter(data_full_dag, type != "Confounder"), aes(label = label), 
    size = 4, nudge_y = -0.46, color = "black", lineheight = 1
  ) +
  geom_dag_label(
    data = filter(data_full_dag, type == "Confounder"), aes(label = label), 
    size = 4, nudge_y = 0.35, color = "black", lineheight = 1
  ) +
  scale_color_manual(
    values = coords |> distinct(type, color) |> pull(color, name = type),
    guide = guide_legend(
      title = NULL, override.aes = list(size = 6), label.position = "bottom"
    )
  ) +
  coord_cartesian(xlim = c(0.98, 3.02)) +
  theme_np_dag() + 
  theme(legend.position = "top")
```

## EDA

### Confounders

```{r}
ggplot(rct, aes(x = social_awareness, y = experiential_learning)) +
  geom_dots(aes(side = ifelse(experiential_learning == 1, "bottom", "top")),
    pch = 19, color = "grey20", scale = 0.2
  ) +
  geom_smooth(
    method = "glm", method.args = list(family = binomial(link = "logit")),
    se = FALSE
  ) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = "Social awareness index", y = "Experiential learning participation")

ggplot(rct, aes(x = age, y = experiential_learning)) +
  geom_dots(aes(side = ifelse(experiential_learning == 1, "bottom", "top")),
    pch = 19, color = "grey20", scale = 0.4
  ) +
  geom_smooth(
    method = "glm", method.args = list(family = binomial(link = "logit")),
    se = FALSE
  ) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = "Age", y = "Experiential learning participation")

ggplot(rct, aes(x = gpa, y = experiential_learning)) +
  geom_dots(aes(side = ifelse(experiential_learning == 1, "bottom", "top")),
    pch = 19, color = "grey20", scale = 0.4
  ) +
  geom_smooth(
    method = "glm", method.args = list(family = binomial(link = "logit")),
    se = FALSE
  ) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = "GPA", y = "Experiential learning participation")

ggplot(rct, aes(x = income, y = experiential_learning)) +
  geom_dots(aes(side = ifelse(experiential_learning == 1, "bottom", "top")),
    pch = 19, color = "grey20", scale = 0.4
  ) +
  geom_smooth(
    method = "glm", method.args = list(family = binomial(link = "logit")),
    se = FALSE
  ) +
  scale_x_continuous(labels = label_dollar()) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = "Income", y = "Experiential learning participation")
```

```{r}
ggplot(rct, aes(x = social_awareness, y = prosocial_index)) +
  geom_point() +
  geom_smooth()

ggplot(rct, aes(x = age, y = prosocial_index)) +
  geom_point() +
  geom_smooth()

ggplot(rct, aes(x = gpa, y = prosocial_index)) +
  geom_point() +
  geom_smooth()

ggplot(rct, aes(x = income, y = prosocial_index)) +
  geom_point() +
  geom_smooth()
```

### Mediator

```{r}
rct |> 
  mutate(experiential_learning_fct = as.logical(experiential_learning)) |> 
  ggplot(aes(x = experiential_learning_fct, y = community_connections)) +
  geom_point(position = position_jitter(height = 0)) +
  geom_boxplot(aes(group = experiential_learning_fct))

ggplot(rct, aes(x = community_connections, y = prosocial_index)) +
  geom_point() +
  labs(x = "Community connections", y = "Prosocial index")
```

### Collider

```{r}
rct |> 
  mutate(experiential_learning_fct = as.logical(experiential_learning)) |> 
  mutate(employed_nonprofit_fct = as.logical(employed_nonprofit)) |> 
  ggplot(aes(x = experiential_learning_fct, y = employed_nonprofit_fct)) +
  geom_point(position = position_jitter())

rct |> 
  mutate(employed_nonprofit_fct = as.logical(employed_nonprofit)) |> 
  ggplot(aes(x = prosocial_index, y = employed_nonprofit_fct)) +
  geom_point(position = position_jitter(width = 0))
```

## Model comparisons

```{r}
model_correct <- lm(
  prosocial_index ~ experiential_learning, 
  data = rct
)

model_adjusted <- lm(
  prosocial_index ~ experiential_learning + social_awareness + age + gpa + income,
  data = rct
)

model_with_mediator <- lm(
  prosocial_index ~ experiential_learning + social_awareness + age + gpa + income + community_connections, 
  data = rct
)

model_collider_selection <- lm(
  prosocial_index ~ experiential_learning + social_awareness + age + gpa + income, 
  data = filter(rct, employed_nonprofit == 1)
)

model_collider_post_treatment <- lm(
  prosocial_index ~ experiential_learning + social_awareness + age + gpa + income + employed_nonprofit, 
  data = rct
)

models <- list(
  "Treatment only" = model_correct,
  "Covariates included" = model_adjusted,
  "Mediator included" = model_with_mediator,
  "Collider (selection)" = model_collider_selection,
  "Collider (post-treatment covariate)" = model_collider_post_treatment
)
```

```{r}
modelsummary(
  models,
  statistic = "[{conf.low}, {conf.high}]",
  fmt = 2,
  gof_map = gof_map,
  add_columns = tibble("True effect" = c(NA, NA, 10)) |> magrittr::set_attr("position", 2)
) |> 
  style_tt(
    i = 3:4, j = 1:7, 
    background = clrs$Prism[6], color = "#ffffff", bold = TRUE
  )
```

```{r}
modelsummary(
  models,
  statistic = "[{conf.low}, {conf.high}]",
  fmt = 2,
  coef_map = c("experiential_learning"),
  gof_map = gof_map,
  add_columns = tibble("True effect" = 10) |> magrittr::set_attr("position", 2)
)
```

```{r}
treatment_effects <- enframe(models) |> 
  mutate(params = map(value, \(x) model_parameters(x))) |> 
  unnest(params) |> 
  filter(Parameter == "experiential_learning") |> 
  mutate(name = factor(name, levels = names(models)))

ggplot(treatment_effects, aes(x = Coefficient, y = fct_rev(name))) +
  geom_vline(xintercept = 10, color = clrs$Prism[2]) +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high)) +
  labs(x = "Treatment effect", y = NULL) +
  theme_np() +
  theme(panel.grid.major.y = element_blank())
```