### 2024-11-21 causal-studies-over-time.qmd
Add NVSQ proportion chart

---
title: Causal work in nonnprofit studies
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "95%",
  warning = FALSE,
  collapse = TRUE,
  cache.extra = 1234  # Change number to invalidate cache
)

options(
  digits = 3, width = 120,
  tinytable_tt_digits = 2
)
```

```{r}
#| label: packages-data
#| warning: false
#| message: false

library(tidyverse)
library(targets)
library(here)

tar_config_set(
  store = here::here('_targets'),
  script = here::here('_targets.R')
)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))

articles <- read_csv(here("data", "manual_data", "Skimming notes_coding.csv"))
nvsq_counts <- read_csv(here("data", "manual_data", "nvsq_counts.csv"))
```

```{r}
#| fig-width: 6
#| fig-height: 4
nvsq_counts_year <- nvsq_counts |>
  filter(Year <= 2020) |>
  group_by(Year) |>
  summarize(total = sum(Count))

plot_data <- articles |>
  filter(year <= 2020) |>
  count(year) |>
  left_join(nvsq_counts_year, by = c("year" = "Year")) |>
  mutate(pct = n / total)

ggplot(plot_data, aes(x = factor(year), y = pct)) +
  geom_col(fill = clrs$Prism[9]) +
  labs(
    x = NULL,
    y = "Percent of NVSQ articles",
    title = "Proportion of causally-focused\narticles in NVSQ, 2010–2020",
    caption = "Article uses an experiment, difference-in-differences, regression discontinuity,\ninstrumental variables, or other method for statistically identifying a causal mechanism\n\n(Voluntas, NML, PAR, JPART, JPAM, and PMR forthcoming)"
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_np() +
  theme(panel.grid.major.x = element_blank())
```