### 2024-11-22 methods/diff-in-diff.qmd
Bigger font for slides

---
title: "Difference-in-differences"
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "95%",
  warning = FALSE,
  collapse = TRUE,
  cache.extra = 1234  # Change number to invalidate cache
)

options(
  digits = 3, width = 120,
  tinytable_tt_digits = 2
)
```

```{r}
#| label: packages-data
#| warning: false
#| message: false

library(tidyverse)
library(targets)
library(modelsummary)
library(tinytable)
library(scales)
library(ggdag)
library(dagitty)
library(parameters)

tar_config_set(
  store = here::here('_targets'),
  script = here::here('_targets.R')
)

did <- tar_read(data_did)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
set_annotation_fonts()

gof_map <- tribble(
  ~raw,            ~clean,               ~fmt, ~omit,
  "nobs",          "N",                  0,    FALSE,
  "r.squared",     "$R^2$",          2,    FALSE,
  "adj.r.squared", "\\(R^2\\) adjusted", 2,    FALSE
)
```

## Parallel trends

```{r}
#| message: false

did |> 
  group_by(school, year) |> 
  summarize(mean = ggplot2::mean_se(prosocial_index, mult = 1.96)) |> 
  unnest(mean) |> 
  ggplot(aes(x = year, y = y, color = school)) +
  geom_vline(xintercept = 2022.5, linetype = "dashed") +
  geom_line(linewidth = 0.25, show.legend = FALSE) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax)) +
  scale_color_manual(values = c(clrs$Prism[2], clrs$Prism[8])) +
  labs(x = NULL, y = "Prosocial index", color = NULL) +
  theme_np(base_size = 16)
```

## Models

```{r}
model_small <- lm(
  prosocial_index ~ school + factor(year) + school*factor(year),
  data = filter(did, year %in% c(2022, 2023))
)

model_small_controls <- lm(
  prosocial_index ~ school + factor(year) + school*factor(year) +
    social_awareness + age + gpa + income,
  data = filter(did, year %in% c(2022, 2023))
)

model_full <- lm(
  prosocial_index ~ school + factor(year) + I(school == "School B" & year >= 2023),
  data = did
)

model_full_controls <- lm(
  prosocial_index ~ school + factor(year) + I(school == "School B" & year >= 2023) +
    social_awareness + age + gpa + income,
  data = did
)

models <- list(
  "Interaction term only" = model_small,
  "Interaction term + covariates" = model_small_controls,
  "Interaction term only (all years)" = model_full,
  "Interaction term + covariates (all years)" = model_full_controls
)
```

```{r}
modelsummary(
  models,
  statistic = "[{conf.low}, {conf.high}]",
  fmt = 2,
  gof_map = gof_map,
  add_columns = tibble("True effect" = c(NA, NA, 10)) |> magrittr::set_attr("position", 2)
)# |> 
  # style_tt(
  #   i = 3:4, j = 1:6, 
  #   background = clrs$Prism[6], color = "#ffffff", bold = TRUE
  # )
```

```{r}
modelsummary(
  models,
  statistic = "[{conf.low}, {conf.high}]",
  fmt = 2,
  coef_map = c(
    `schoolSchool B:factor(year)2023` = "School B × 2023",
    `I(school == "School B" & year >= 2023)TRUE` = "School B × 2023"
  ),
  gof_map = gof_map,
  add_columns = tibble("True effect" = 10) |> magrittr::set_attr("position", 2)
)
```

```{r}
treatment_effects <- enframe(models) |> 
  mutate(params = map(value, \(x) model_parameters(x))) |> 
  unnest(params) |> 
  filter(Parameter %in% c("schoolSchool B:factor(year)2023", 'I(school == "School B" & year >= 2023)TRUE')) |> 
  mutate(name = factor(name, levels = names(models)))

ggplot(treatment_effects, aes(x = Coefficient, y = fct_rev(name))) +
  geom_vline(xintercept = 10, color = clrs$Prism[2]) +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high)) +
  labs(x = "Treatment effect", y = NULL) +
  theme_np(base_size = 16) +
  theme(panel.grid.major.y = element_blank())
```