### 2024-11-22 methods/rdd-sharp.qmd

---
title: "Sharp RDD"
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "95%",
  warning = FALSE,
  collapse = TRUE,
  cache.extra = 1234  # Change number to invalidate cache
)

options(
  digits = 3, width = 120,
  tinytable_tt_digits = 2
)
```

```{r}
#| label: packages-data
#| warning: false
#| message: false

library(tidyverse)
library(targets)
library(modelsummary)
library(tinytable)
library(scales)
library(ggdag)
library(dagitty)
library(parameters)
library(rdrobust)
library(rddensity)

tar_config_set(
  store = here::here('_targets'),
  script = here::here('_targets.R')
)

rdd <- tar_read(data_rdd_sharp)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
set_annotation_fonts()

gof_map <- tribble(
  ~raw,            ~clean,               ~fmt, ~omit,
  "nobs",          "N",                  0,    FALSE,
  "r.squared",     "$R^2$",          2,    FALSE,
  "adj.r.squared", "\\(R^2\\) adjusted", 2,    FALSE
)
```

```{r}
#| include: false
#' Tidy a `rdrobust` regression discontinuity design object
#'
#' @param x A `rdrobust` object
#' @param ... extra arguments (not used)
#' @examples
#' \dontrun{
#' x <- runif(1000,-1,1)
#' y <- 3 * x + 2 * (x>=0) + rnorm(1000)
#' mod <- rdrobust(y,x)
#' 
#' tidy(mod)
#' 
#' glance(mod)
#' }
#' @export
tidy.rdrobust <- function(object, ...){
    ret <- tibble(term = row.names(object$coef), 
                  estimate = object$coef[, 1], 
                  std.error = object$se[, 1], 
                  statistic = object$z[, 1],
                  p.value = object$pv[, 1], 
                  conf.low = object$ci[,1],
                  conf.high = object$ci[, 2])
    ret
}

#' Glance a `rdrobust` regression discontinuity design object
#'
#' @inheritParams tidy.rdrobust
#' @export
glance.rdrobust <- function(object, ...){
    ret <- tibble(nobs.left = object$N[1],
                  nobs.right = object$N[2],
                  nobs.effective.left = object$N_h[1],
                  nobs.effective.right = object$N_h[2],
                  cutoff = object$c,
                  order.regression = object$q,
                  order.bias = object$q,
                  kernel = object$kernel,
                  bwselect = object$bwselect)
    ret
}
```

```{r}
rdd |> 
  ggplot(aes(
    x = entrance_score, y = as.logical(experiential_learning), 
    color = factor(experiential_learning)
  )) +
  geom_point(
    size = 0.5, alpha = 0.5,
    position = position_jitter(width = 0, height = 0.25, seed = 1234)
  ) +
  geom_vline(xintercept = 75) +
  labs(
    x = "Entrance test score", 
    y = "Participated in\nexperiential learning program"
  ) +
  scale_color_manual(values = c(clrs$Prism[2], clrs$Prism[8]), guide = "none") +
  theme_np()
```

```{r}
ggplot(rdd, aes(x = entrance_score, fill = as.logical(experiential_learning))) +
  geom_histogram(binwidth = 1, color = "white", boundary = 0) +
  geom_vline(xintercept = 75) +
  labs(x = "Entrance exam score", y = "Count", fill = "In program") +
  scale_fill_manual(values = c(clrs$Prism[2], clrs$Prism[8])) +
  theme_np()
```

```{r}
test_density <- rddensity(rdd$entrance_score, c = 75)
# summary(test_density)

plot_density_test <- rdplotdensity(
  rdd = test_density,
  X = rdd$entrance_score,
  type = "both"
)
```

```{r}
ggplot(rdd, aes(x = entrance_score, y = prosocial_index, color = as.logical(experiential_learning))) +
  geom_point(size = 0.5, alpha = 0.5) +
  geom_smooth(data = filter(rdd, entrance_score <= 75, entrance_score >= 70),
              method = "lm", se = FALSE, linewidth = 2) +
  geom_smooth(data = filter(rdd, entrance_score > 75, entrance_score <= 80),
              method = "lm", se = FALSE, linewidth = 2) +
  geom_vline(xintercept = 75.5) +
  coord_cartesian(xlim = c(65, 85), ylim = c(60, 100)) +
  labs(x = "Entrance test score", y = "Prosocial index", color = "Participated") +
  scale_color_manual(values = c(clrs$Prism[2], clrs$Prism[8])) +
  theme_np(base_size = 16)
```

```{r}
blah <- rdplot(y = rdd$prosocial_index, x = rdd$entrance_score, c = 75, hide = TRUE)
blah$rdplot +
  theme_np() +
  labs(title = NULL, x = "Entrance test score", y = "Prosocial index")
```

```{r}
#| warning: false
model_rdrobust <- rdrobust(y = rdd$prosocial_index, x = rdd$entrance_score, c = 75)

model_rdrobust |>
  summary()

tidy(model_rdrobust)
```

```{r}
model_bw5 <- lm(
  prosocial_index ~ entrance_score + experiential_learning,
  data = filter(rdd, entrance_score >= (75 - 5) & entrance_score <= (75 + 5))
)

model_bw10 <- lm(
  prosocial_index ~ entrance_score + experiential_learning,
  data = filter(rdd, entrance_score >= (75 - 10) & entrance_score <= (75 + 10))
)

models <- list(
  "BW = 5" = model_bw5,
  "BW = 10" = model_bw10,
  "rdrobust" = model_rdrobust
)

modelsummary(
  models,
  statistic = "[{conf.low}, {conf.high}]",
  fmt = 2,
  coef_map = c(
    `experiential_learning` = "Experiential learning",
    `Conventional` = "Experiential learning"
  ),
  gof_map = gof_map,
  add_columns = tibble("True effect" = 8.5) |> magrittr::set_attr("position", 2)
)
```

```{r}
treatment_effects <- enframe(models) |> 
  mutate(params = map(value, \(x) tidy(x, conf.int = TRUE))) |> 
  unnest(params) |> 
  filter(term %in% c("experiential_learning", "Conventional")) |> 
  mutate(name = factor(name, levels = names(models)))

ggplot(treatment_effects, aes(x = estimate, y = fct_rev(name))) +
  geom_vline(xintercept = 8.5, color = clrs$Prism[2]) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  labs(x = "Treatment effect", y = NULL) +
  theme_np(base_size = 16) +
  theme(panel.grid.major.y = element_blank())
```