### 2024-11-22 presentations/2024-11-22_arnova/index.qmd
Add embedded google presentation

---
title: "ARNOVA"
date: 2024-11-22
toc: false
resources: slides
abstract: |
  Annual conference of the Association for Research on Nonprofit Organizations and Voluntary Action (ARNOVA); Washington, D.C.
abstract-title: "Location"
published-title: "Presented"
image: first-slide.png
page-layout: full
---

<div class="ratio ratio-16x9">
<iframe class="slide-deck" src="https://docs.google.com/presentation/d/e/2PACX-1vRTS3Uve-PLfIOJjtIFIQgMFz1Zs3HhArdtNvgFd6VGnA_kuxqrRelNHHE2Tv4CFRty77ZB5EdkNRon/embed?start=false&loop=false&delayms=3000" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>
</div>