### 2024-11-22 methods/iv.qmd
Add IV randomized promotion data

---
title: "Instrumental variables"
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 6 * 0.618,
  fig.retina = 3,
  dev = "ragg_png",
  fig.align = "center",
  out.width = "95%",
  warning = FALSE,
  collapse = TRUE,
  cache.extra = 1234  # Change number to invalidate cache
)

options(
  digits = 3, width = 120,
  tinytable_tt_digits = 2
)
```

```{r}
#| label: packages-data
#| warning: false
#| message: false

library(tidyverse)
library(targets)
library(modelsummary)
library(tinytable)
library(scales)
library(ggdag)
library(dagitty)
library(parameters)
library(estimatr)

tar_config_set(
  store = here::here('_targets'),
  script = here::here('_targets.R')
)

iv <- tar_read(data_iv)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
set_annotation_fonts()

gof_map <- tribble(
  ~raw,            ~clean,               ~fmt, ~omit,
  "nobs",          "N",                  0,    FALSE,
  "r.squared",     "$R^2$",          2,    FALSE,
  "adj.r.squared", "\\(R^2\\) adjusted", 2,    FALSE
)
```

```{r}
model_iv_basic <- iv_robust(
  prosocial_index ~ experiential_learning | promoted, 
  data = iv
)

model_iv_controls <- iv_robust(
  prosocial_index ~ experiential_learning + social_awareness + age + gpa + income | 
    promoted + social_awareness + age + gpa + income, 
  data = iv
)

model_naive <- lm(
  prosocial_index ~ experiential_learning, 
  data = iv
)

model_naive_controls <- lm(
  prosocial_index ~ experiential_learning + social_awareness + age + gpa + income, 
  data = iv
)

models <- list(
  "IV" = model_iv_basic,
  "IV with controls" = model_iv_controls,
  "Naive OLS" = model_naive,
  "OLS with controls" = model_naive_controls
)
```

```{r}
modelsummary(
  models,
  statistic = "[{conf.low}, {conf.high}]",
  fmt = 2,
  coef_map = c("experiential_learning"),
  gof_map = gof_map,
  add_columns = tibble("True effect" = 10) |> magrittr::set_attr("position", 2)
)
```

```{r}
treatment_effects <- enframe(models) |> 
  mutate(params = map(value, \(x) model_parameters(x))) |> 
  unnest(params) |> 
  filter(Parameter == "experiential_learning") |> 
  mutate(name = factor(name, levels = names(models)))

ggplot(treatment_effects, aes(x = Coefficient, y = fct_rev(name))) +
  geom_vline(xintercept = 10, color = clrs$Prism[2]) +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high)) +
  labs(x = "Treatment effect", y = NULL) +
  theme_np(base_size = 16) +
  theme(panel.grid.major.y = element_blank())
```